Measuring directionality in double-beta decay and neutrino interactions with kiloton-scale scintillation detectors

### 2014-01-14 Direction_JINST_v11.tex
the last few versions. --Lindley

The advantage of liquid scintillators for measurements in the
$\sim$1~MeV range is their scalability from 1~ton to 1~kiloton while
providing energy resolutions of $\sigma(E)=\sim$5 \% $/\sqrt{E(MeV)}$\cite{kam2013,borexino}. This is roughly a factor of
two better than water Cherenkov detectors, the other developed
technology for neutrino detectors that can be economically scaled to these large
masses. However, this energy resolution is much poorer than other technologies being used for $0\nu\beta\beta$ searches: Ge detectors\cite{gerda2013}, Te bolometers\cite{Alessandria:2011rc}, tracking detectors\cite{SuperNEMO}, liquid Xe time projection chambers (TPCs)\cite{EXO2012} and high pressure gaseous Xe TPCs\cite{NEXTsipm}.

Scintillation light is isotropic. At these low energies, it does not contain sufficient information to reconstruct the track of the outgoing particles, although at higher energies it may\cite{john}.  Cherenkov light is also
produced for electrons above threshold. Most is absorbed and re-emitted as part of the
scintillation processes; however some fraction retains its
directional information. If this directional Cherenkov light can be
isolated from the copious isotropic scintillation light, it may be
possible to reconstruct the direction of the primary particle. The addition of directionality is a powerful tool
for background rejection, especially for $0\nu\beta\beta$ searches. In the high pressure TPCs, reduction factors on the order of $\sim10^4$ have been achieved\cite{Gotthard}. It is also possible to look for new physics in the angular correlation of the emerging electrons\cite{newphysics0nuBB}, as has been proposed for the tracking-based detectors\cite{SuperNEMO,NEXTsipm}. The addition of a directional signal would make large-scale liquid scintillator detectors more competitive for the next generation of $0\nu\beta\beta$ searches and merits investigation.

This is the first in a series of papers exploring directionality in large-scale liquid scintillator detectors. In this paper, we develop a technique for
separating the Cherenkov and scintillation light using the photon arrival
times and evaluate several detector advances in timing, photodetector 
spectral response, and scintillator emission spectra that would allow
the realization of direction reconstruction in kilo-ton scale
scintillating neutrino detectors. This is a different from the direction reconstruction described for high-energy neutrino interactions\cite{john} or that for neutrons from inverse beta decay\cite{chooz,dcDirection}.  We then use these results as input into a traditional direction reconstruction developed for water Cherenkov detectors. Since the reconstruction of the direction of $\sim$1~MeV particles has not been achieved before, we start these studies with the most simple case of a single particle at the center of the detector. We also start with an easier test case of a 5~MeV electron like that from a $^{8}$B solar neutrino interaction. With the higher photon statistics at this energy, we verify the technique and the different detector parameters that affect it. We then study the technique for two lower energies, 1.4 and 2.1 MeV, that are more relevant to $0\nu\beta\beta$.

@@

The Cherenkov light produced at wavelengths shorter than the
absorption cutoff of the scintillator will be absorbed and re-emitted
as isotropic light, but wavelengths longer than this cutoff will
propagate across the detector, retaining their directional
information. The absorption cutoff for the example scintillator above is 370~nm. The index of refraction at  n(370nm)=1.466 and can be translated into an effective Cherenkov threshold for electrons at 0.188~MeV. For a 5~MeV electron, this yields 685 Cherenkov photons per event when integrating from the cutoff wavelength at 370~nm to 550~nm,  the wavelength above which a small number of photons is detected due to a standard bialkali photocathode's quantum efficiency. Lowering the energy to 1~MeV yields 82 Cherenkov photons between 370-550~nm.

@@

In neutrino-electron scattering events, a single electron emerges with a distribution of energies with a maximum energy related to the incoming neutrino's energy. In comparison, $0\nu\beta\beta$ events are more complicated with two electron emerge with a combined energy equal
to the Q-value of the particular isotope. The individual
electrons follow distributions of energies and angular correlations which depend on the underlying $0\nu\beta\beta$ decay mechanism\cite{SuperNEMO,newphysics0nuBB,bandv}. Figure 1 shows a simulated $^{116}$Cd $0\nu\beta\beta$ event in a model with light Majorana neutrino exchange, for which a probable case is the emission of two electrons with comparable energies at a large angle relative to each other.

High Q-value candidates are preferred for $0\nu\beta\beta$  for two reasons. First, the measured half-life is inversely proportional to the phase-space factor, therefore the measured rate is expected to be higher from isotopes with higher Q-values. Second, the main background for these experiments come from the daughters of the $^{238}$U and $^{232}$Th decay chains. The Compton shoulder from gamma-rays is particularly problematic. The 2.6~MeV gamma-ray from $^{208}$Tl decay is the highest energy gamma from these decay chains and therefore isotopes above 2.6~MeV are preferred.  Experiments using isotopes with Q-values below this energy must compensate with improved background rejection techniques.

Most of the high Q-value candidates\cite{tabledbb}
have been considered as a dopant for a liquid scintillator:
$^{150}$Nd (Q=3.367~MeV)\cite{minfang,nd1}, $^{96}$Zr (Q=3.350~MeV)\cite{zr1},
$^{100}$Mo (Q=3.034~MeV)\cite{mo1}, $^{82}$Se (Q=2.995~MeV)\cite{qdot},
$^{116}$Cd (Q=2.81~MeV)\cite{qdot, cd1}, $^{130}$Te (Q=2.533~MeV)\cite{qdot, biller},
$^{136}$Xe (Q=2.479~MeV)\cite{KZ0nu} and $^{124}$Sn (Q=2.29~MeV)\cite{sn1}.
Xenon gas readily dissolves into liquid scintillator. For the other isotopes,
a suitable organometallic compound needs to be found that produces a stable
scintillator with a long attenuation length in the wavelength region of
interest. Recently nanocrystals formed by candidate isotopes have been explored as an alternative to doping by single atoms,\cite{qdot,qdot2}.

@@

The detector geometry is a sphere of 6.5~m radius filled with
scintillator. Figure \ref{detector_view} shows the geometry and the
Cherenkov light from an example $^{116}$Cd $0\nu\beta\beta$ event. The
default scintillator properties have been chosen to match a KamLAND-like
scintillator\cite{kamland2003}: 80\% n-dodecane, 20\% pseudocumene and 1.52~g/l PPO. The
scintillator properties implemented in the simulation include the
atomic composition and density ($\rho$ = 0.78~g/ml), the
wavelength-dependent attenuation length\cite{tajimaMaster} and
refractive index\cite{OlegThesis}, the scintillation emission
spectrum\cite{tajimaMaster}, emission rise time ($\tau_r$ = 1.0~ns)
and emission decay time constants ($\tau_{d1}$ = 6.9~ns and
$\tau_{d2}$ = 8.8~ns with relative weights of 0.87 and 
0.13)\cite{tajimaThesis}, scintillator light yield (9030 photons/MeV)
and the Birks constant ($kB$ $\approx$ 0.1~mm/MeV)\cite{ChrisThesis}.  This is a standard scintillator. The attenuation length at 400~nm, the position of the peak standard bialkali photocathode efficiency, is 25~m. The attenuation length drops precipitously between 370~nm and 360~nm from 6.5~m to 0.65~m. We use this drop to define the cutoff wavelength at 370~nm. Variations from the baseline KamLAND case are discussed below.

@@

\begin{figure}
        \begin{center}
        \includegraphics[scale=0.4]{graphs2/PEtime_Cerenkov_scattered_vs_unscattered.pdf}
        \caption[]{Cherenkov photoelectron (PE) arrival times after application
        of the transit-time spread (TTS) for the simulation
        of 1000 electrons (5~MeV). The default simulation is shown with optical scattering turned on: all photons (black), un-scattered photons (red dotted), and scattered (blue dashed). Scattering causes a very small tail at longer times as expected. \label{scattplot}}
        \end{center}
\end{figure}

@@

However, advances in the production of commercial quantum dot samples
could yield quantum dots which have similar, single peak emission
shapes at shorter wavelengths. This case has been simulated using the
same spectral shape of the measured core-shell quantum dot emission
but shifted to shorter wavelengths such that the emission peak is
centered at 384~nm. This peak emission value has been measured for
other types of QDs, however with a much more pronounced 
tail\cite{qdot2}. The resulting PE time distribution shows improved
separation of Cherenkov and scintillation light compared to the
default simulation. After the 34.0~ns cut on the TTS-smeared PE time
we obtain a Cherenkov/scintillation ratio of $R_{C/S}=0.86$ (107 PE
from Cherenkov light and 124 PE from scintillation). The number of
Cherenkov-induced PEs after the time cut is unchanged while the number
of PEs coming from scintillation light is decreased due to the higher
average photon travel times.

\begin{figure}[tbh]
        \begin{center}
        %%\subfigure{\includegraphics[scale=0.295]{graphs/hDirX_v1.pdf}}
        %%\subfigure{\includegraphics[scale=0.295]{graphs/hDirY_v1.pdf}}
        %%\subfigure{\includegraphics[scale=0.295]{graphs/hDirZ_v1.pdf}}
        %%\subfigure{\includegraphics[scale=0.295]{graphs/hVtxX_v1.pdf}}
        %%\subfigure{\includegraphics[scale=0.295]{graphs/hVtxY_v1.pdf}}
        %%\subfigure{\includegraphics[scale=0.295]{graphs/hVtxZ_v1.pdf}}
        \subfigure{\includegraphics[scale=0.375]{graphs2/hDir_pq2_5p0.pdf}}
        \subfigure{\includegraphics[scale=0.375]{graphs2/hVtx_pq2_5p0.pdf}}
        \subfigure{\includegraphics[scale=0.375]{graphs2/hDir_pq2_2p1.pdf}}
        \subfigure{\includegraphics[scale=0.375]{graphs2/hVtx_pq2_2p1.pdf}}
        \subfigure{\includegraphics[scale=0.375]{graphs2/hDir_pq2_1p4.pdf}}
        \subfigure{\includegraphics[scale=0.375]{graphs2/hVtx_pq2_1p4.pdf}}
        \caption[]{\label{fig:reco} (Left) The reconstructed direction,
        $(p_x/|\vec{p}|, p_y/|\vec{p}|,
        p_z/|\vec{p}|)$, for the simulation of 1000 electrons. In the simulation the electrons are produced along the
        x-axis, $\vec{p}/|\vec{p}|$ = (1,0,0), and originate
        from the center of the 6.5m-radius detector, $\vec{r}$ =
        (0,0,0). Only photons with arrival time of $t<$ 34.0~ns are used
        in the reconstruction. The quantum efficiency of the bialkali
        photocathode is taken into account. (Right) The reconstructed
        vertex position, $(x,y,z)$, for the same simulation. From Top to Bottom, 5~MeV, 2.1~MeV and 1.4 MeV are shown.}
\end{center}
\end{figure}

@@

To start, we are neglecting the effects of position dependence. In future work the arrival times can be corrected by the time of flight from the reconstructed vertex and the position and direction fitted simultaneously. For isotropic light, the vertex reconstruction uncertainty leads to an additional smearing of the arrival time distribution of 0.15~ns for every 3~cm of position reconstruction uncertainty. For directional light, the vertex reconstruction uncertainty leads to an effective shift of the arrival time distribution. This would change the ideal time cut by 0.15~ns  for every 3~cm of position reconstruction uncertainty. We have studied other time cuts from 33~ns to 34.5~ns and the reconstruction remained reasonable.

@@

The two additional simulation sets with 1.4 MeV and 2.1 MeV electrons were
generated using the default simulation configuration described in section
\ref{sim_section}. The PE time distribution for the default settings is shown
in figure \ref{time_plots_comparison} (a) for 5~MeV electrons. The shapes
of the scintillation and Cherenkov spectra are similar for the lower energies (not
shown here). In figure \ref{Edep_NPE}, the energy-dependent mean number of
PEs per event after the 34.0~ns time cut is shown for Cherenkov-induced and
scintillation-induced PEs, as well as their ratio $R_{C/S}$. The mean number
of PEs from Cherenkov (scintillation) light for electron energies 1.4(2.1)~MeV is 21.8 (52.1), 38.4 (76.8)
and 108 (171) for 1.4~MeV, 2.1~MeV and 5~MeV, respectively. This gives the
ratios $R_{C/S}$ = 0.42, 0.50 and 0.63: The decrease in Cherenkov-induced
PEs is stronger than the decrease in scintillation-induced PEs as the energy
is lowered.

@@

We have developed a technique to separate scintillation and Cherenkov light to reconstruct direction of electrons with energies 5~MeV, 2.1~MeV and 1.4~MeV in a liquid scintillator detector. These energies have been chosen to represent typical neutrino-electron elastic scattering energies and
$0\nu\beta\beta$ energies. The Cherenkov threshold for an electron in a typical liquid scintillator is $\sim$0.188~MeV.

While scintillation light is isotropic, Cherenkov light with wavelengths above the absorption cutoff $\sim$370~nm carries the information about the direction of the electrons. All light shorter than this gets absorbed and re-emitted isotropically as a part of the scintillation process. On average scintillation light is delayed with respect to the direct Cherenkov light due to chromatic dispersion and the finite time of the scintillation processes, therefore early light contains directional information.

Using a Geant4~simulation of a spherical detector with radius of 6.5~m and photodetectors with TTS of 0.1~ns, we have showed that for electrons originated at the center a time cut on the early light was effective at isolating the directional light, improving the ratio of Cherenkov light to scintillation light from 0.02 to 0.63. This ratio is degraded by a factor of 2.5 if current photodetectors with $\sim$1~ns resolution were used. This can be improved by factors of 1.6 or 1.4 if more red sensitive photodetectors or scintillators with narrower emission spectra are used.

Reconstruction algorithms developed for water Cherenkov detectors have been applied to this early light. The algorithms are able to converge on reasonable reconstructed vertices and directions for all simulated energies: 5~MeV, 2.1~MeV, 1.4~MeV. As expected, the reduction in photon statistics with lower energies leads to a broadening of the reconstructed vertex by $\sim$1~cm. Similarly, the mean reconstructed direction drops from 0.88 to 0.74 as the $\cos(\theta)$ distribution broadens.

We have shown that this technique is promising, and we plan to continue work on the topic. The ability to reconstruct direction in kiloton-scale scintillation detectors would expand capabilities of neutrino experiments, especially those also searching for neutrino-less double-beta decay. More generally, this technique could be applied wherever scintillation-based detectors are used.

>>>>

The advantage of liquid scintillators for measurements in the
$\sim$1~MeV range is their scalability from 1~ton to 1~kiloton while
providing energy resolutions of $\sigma(E)=\sim$5 \% $/\sqrt{E(MeV)}$\cite{kam2013,borexino}. This is roughly a factor of
two better than for water Cherenkov detectors, the other developed
technology for neutrino detectors that can be economically scaled to these large
masses. However, this energy resolution is much poorer than other technologies being used for $0\nu\beta\beta$ searches: Ge detectors\cite{gerda2013}, Te bolometers\cite{Alessandria:2011rc}, tracking detectors\cite{SuperNEMO}, liquid Xe time projection chambers (TPCs)\cite{EXO2012} and high pressure gaseous Xe TPCs\cite{NEXTsipm}.

Scintillation light is isotropic. At these low energies, it does not contain sufficient information to reconstruct the track of the outgoing particles, although at higher energies it may\cite{john}.  Cherenkov light is also
produced for electrons above threshold. Most is absorbed and re-emitted as part of the
scintillation processes; however some fraction retains its
directional information. If this directional Cherenkov light can be
isolated from the copious isotropic scintillation light, it may be
possible to reconstruct the direction of the primary particle. The addition of directionality is a powerful tool
for background rejection, especially for $0\nu\beta\beta$ searches. In high pressure TPCs, reduction factors on the order of $\sim10^4$ have been achieved\cite{Gotthard}. It is also possible to look for new physics in the angular correlation of the emerging electrons\cite{newphysics0nuBB}, as has been proposed for the tracking-based detectors\cite{SuperNEMO,NEXTsipm}. The addition of a directional signal would make large-scale liquid scintillator detectors more competitive for the next generation of $0\nu\beta\beta$ searches. 
%and merits investigation.

This is the first in a series of papers exploring directionality in large-scale liquid scintillator detectors. In this paper, we develop a technique for
separating the Cherenkov and scintillation light using the photon arrival
times and evaluate several detector advances in timing, photodetector 
spectral response, and scintillator emission spectra that would allow
the realization of direction reconstruction in kilo-ton scale
scintillating neutrino detectors. This is different from the direction reconstruction described for high-energy neutrino interactions\cite{john} or that for neutrons from inverse beta decay\cite{chooz,dcDirection}.  We then use these results as input into a traditional direction reconstruction developed for water Cherenkov detectors. Since the reconstruction of the direction of $\sim$1~MeV particles has not been achieved before, we start these studies with the %most
simple case of a single particle at the center of the detector. We also start with an easier test case of a 5~MeV electron like that from a $^{8}$B solar neutrino interaction. With the higher photon statistics at this energy, we verify the technique and the different detector parameters that affect it. We then study the technique for two lower energies, 1.4 and 2.1 MeV, that are more relevant to $0\nu\beta\beta$.

@@

The Cherenkov light produced at wavelengths shorter than the
absorption cutoff of the scintillator will be absorbed and re-emitted
as isotropic light, but wavelengths longer than this cutoff will
propagate across the detector, retaining their directional
information. The absorption cutoff for the example scintillator above is 370~nm. The index of refraction at $\lambda$=370~nm is n=1.466, which translates
%n(370nm)=1.466 and can be translated 
into an effective Cherenkov threshold for electrons at 0.188~MeV. For a 5~MeV electron, this yields 685 Cherenkov photons per event when integrating from the cutoff wavelength at 370~nm to 550~nm,  the wavelength above which a small number of photons is detected due to a standard bialkali photocathode's quantum efficiency. Lowering the energy to 1~MeV yields 82 Cherenkov photons between 370-550~nm.

@@

In neutrino-electron scattering events, a single electron emerges with a distribution of energies with a maximum energy related to the incoming neutrino's energy. In comparison, $0\nu\beta\beta$ events are more complicated, with two electrons emerging with a combined energy equal
to the Q-value of the particular isotope. The individual
electrons follow distributions of energies and angular correlations which depend on the underlying $0\nu\beta\beta$ decay mechanism\cite{SuperNEMO,newphysics0nuBB,bandv}. Figure 1 shows a simulated $^{116}$Cd $0\nu\beta\beta$ event in a model with light Majorana neutrino exchange, for which a probable case is the emission of two electrons with comparable energies at a large angle relative to each other.

High Q-value candidates are preferred for $0\nu\beta\beta$  for two reasons. First, as the measured half-life is inversely proportional to the phase-space factor, the measured rate is expected to be higher from isotopes with higher Q-values. Second, the main background for these experiments come from the daughters of the $^{238}$U and $^{232}$Th decay chains. The Compton shoulder from gamma-rays is particularly problematic. The 2.6~MeV gamma-ray from $^{208}$Tl decay is the highest energy gamma from these decay chains and so isotopes above 2.6~MeV are preferred.  Experiments using isotopes with Q-values below this energy must compensate with improved background rejection techniques.

Most of the high Q-value candidates\cite{tabledbb}
have been considered as a dopant for a liquid scintillator:
$^{150}$Nd (Q=3.367~MeV)\cite{minfang,nd1}, $^{96}$Zr (Q=3.350~MeV)\cite{zr1},
$^{100}$Mo (Q=3.034~MeV)\cite{mo1}, $^{82}$Se (Q=2.995~MeV)\cite{qdot},
$^{116}$Cd (Q=2.81~MeV)\cite{qdot, cd1}, $^{130}$Te (Q=2.533~MeV)\cite{qdot, biller},
$^{136}$Xe (Q=2.479~MeV)\cite{KZ0nu} and $^{124}$Sn (Q=2.29~MeV)\cite{sn1}.
Xenon gas readily dissolves into liquid scintillator. For the other isotopes,
a suitable organometallic compound needs to be found that produces a stable
scintillator with a long attenuation length in the wavelength region of
interest. Recently nanocrystals formed by candidate isotopes have been explored as an alternative to doping by single atoms\cite{qdot,qdot2}.

@@

The detector geometry is a sphere of 6.5~m radius filled with
scintillator. Figure \ref{detector_view} shows the geometry and the
Cherenkov light from an example $^{116}$Cd $0\nu\beta\beta$ event. The
default scintillator properties have been chosen to match a KamLAND-like
scintillator\cite{kamland2003}: 80\% n-dodecane, 20\% pseudocumene and 1.52~g/l PPO. The
scintillator properties implemented in the simulation include the
atomic composition and density ($\rho$ = 0.78~g/ml), the
wavelength-dependent attenuation length\cite{tajimaMaster} and
refractive index\cite{OlegThesis}, the scintillation emission
spectrum\cite{tajimaMaster}, emission rise time ($\tau_r$ = 1.0~ns)
and emission decay time constants ($\tau_{d1}$ = 6.9~ns and
$\tau_{d2}$ = 8.8~ns with relative weights of 0.87 and 
0.13)\cite{tajimaThesis}, scintillator light yield (9030 photons/MeV),
and the Birks constant ($kB$ $\approx$ 0.1~mm/MeV)\cite{ChrisThesis}.  This is a standard scintillator. The attenuation length at 400~nm, the position of the peak standard bialkali photocathode efficiency, is 25~m. The attenuation length drops precipitously between 370~nm and 360~nm from 6.5~m to 0.65~m. We use this drop to define the cutoff wavelength at 370~nm. Variations from the baseline KamLAND case are discussed below.

@@

\begin{figure}
        \begin{center}
        \includegraphics[scale=0.4]{graphs2/PEtime_Cerenkov_scattered_vs_unscattered.pdf}
        \caption[]{Cherenkov photoelectron (PE) arrival times after application
        of the transit-time spread (TTS) for the simulation
        of 1000 electrons (5~MeV). The default simulation is shown with optical scattering turned on: all photons (black solid), un-scattered photons (red dotted), and scattered (blue dashed). Scattering causes a very small tail at longer times as expected. \label{scattplot}}
        \end{center}
\end{figure}

@@

However, advances in the production of commercial quantum dot samples
could yield quantum dots that have similar, single peak emission
shapes at shorter wavelengths. This case has been simulated using the
same spectral shape of the measured core-shell quantum dot emission
but shifted to shorter wavelengths such that the emission peak is
centered at 384~nm. This peak emission value has been measured for
other types of QDs, however with a much more pronounced 
tail\cite{qdot2}. The resulting PE time distribution shows improved
separation of Cherenkov and scintillation light compared to the
default simulation. After the 34.0~ns cut on the TTS-smeared PE time
we obtain a Cherenkov/scintillation ratio of $R_{C/S}=0.86$ (107 PE
from Cherenkov light and 124 PE from scintillation). The number of
Cherenkov-induced PEs after the time cut is unchanged while the number
of PEs coming from scintillation light is decreased due to the higher
average photon travel times.

\begin{figure}[tbh]
        \begin{center}
        %%\subfigure{\includegraphics[scale=0.295]{graphs/hDirX_v1.pdf}}
        %%\subfigure{\includegraphics[scale=0.295]{graphs/hDirY_v1.pdf}}
        %%\subfigure{\includegraphics[scale=0.295]{graphs/hDirZ_v1.pdf}}
        %%\subfigure{\includegraphics[scale=0.295]{graphs/hVtxX_v1.pdf}}
        %%\subfigure{\includegraphics[scale=0.295]{graphs/hVtxY_v1.pdf}}
        %%\subfigure{\includegraphics[scale=0.295]{graphs/hVtxZ_v1.pdf}}
        \subfigure{\includegraphics[scale=0.375]{graphs2/hDir_pq2_5p0.pdf}}
        \subfigure{\includegraphics[scale=0.375]{graphs2/hVtx_pq2_5p0.pdf}}
        \subfigure{\includegraphics[scale=0.375]{graphs2/hDir_pq2_2p1.pdf}}
        \subfigure{\includegraphics[scale=0.375]{graphs2/hVtx_pq2_2p1.pdf}}
        \subfigure{\includegraphics[scale=0.375]{graphs2/hDir_pq2_1p4.pdf}}
        \subfigure{\includegraphics[scale=0.375]{graphs2/hVtx_pq2_1p4.pdf}}
        \caption[]{\label{fig:reco} (Left) The reconstructed direction,
        $(p_x/|\vec{p}|, p_y/|\vec{p}|,
        p_z/|\vec{p}|)$, for the simulation of 1000 electrons. In the simulation the electrons are produced along the
        x-axis, $\vec{p}/|\vec{p}|$ = (1,0,0), and originate
        at the center of the 6.5m-radius detector, $\vec{r}$ =
        (0,0,0). Only photons with arrival time of $t<$ 34.0~ns are used
        in the reconstruction. The quantum efficiency of the bialkali
        photocathode is taken into account. (Right) The reconstructed
        vertex position, $(x,y,z)$, for the same simulation. From Top to Bottom, 5~MeV, 2.1~MeV and 1.4 MeV are shown.}
\end{center}
\end{figure}

@@

To start, we are neglecting the effects of position dependence. In future work the arrival times can be corrected by the time of flight from the reconstructed vertex and the position and direction fitted simultaneously. For isotropic light, the vertex reconstruction uncertainty leads to an additional smearing of the arrival time distribution of 0.15~ns for every 3~cm of position reconstruction uncertainty. For directional light, the vertex reconstruction uncertainty leads to an effective shift of the arrival time distribution. This would change the ideal time cut by 0.15~ns  for every 3~cm of position reconstruction uncertainty. We have studied other time cuts from 33~ns to 34.5~ns and the reconstruction remains reasonable.

@@

The two additional simulation sets with 1.4 MeV and 2.1 MeV electrons were
generated using the default simulation configuration described in section
\ref{sim_section}. The PE time distribution for the default settings is shown
in figure \ref{time_plots_comparison} (a) for 5~MeV electrons. The shapes
of the scintillation and Cherenkov spectra are similar for the lower energies (not
shown here). In figure \ref{Edep_NPE}, the energy-dependent mean number of
PEs per event after the 34.0~ns time cut is shown for Cherenkov-induced and
scintillation-induced PEs, as well as their ratio $R_{C/S}$. The mean number
of PEs from Cherenkov (scintillation) light for electron energies of 1.4, 2.1, and 5~MeV is 21.8 (52.1), 38.4 (76.8)
and 108 (171) for 1.4~MeV, 2.1~MeV and 5~MeV, respectively. This gives the
ratios $R_{C/S}$ = 0.42, 0.50 and 0.63: The decrease in Cherenkov-induced
PEs is stronger than the decrease in scintillation-induced PEs as the energy
is lowered.

@@

We have developed a technique to separate scintillation and Cherenkov light to reconstruct direction of electrons with energies 5~MeV, 2.1~MeV and 1.4~MeV in a liquid scintillator detector. These energies have been chosen to represent typical neutrino-electron elastic scattering energies and
$0\nu\beta\beta$ energies. The Cherenkov threshold for an electron in a typical liquid scintillator is $\sim$0.2~MeV.

While scintillation light is isotropic, Cherenkov light with wavelengths above the absorption cutoff $\sim$370~nm carries the information about the direction of the electrons. All light shorter than this gets absorbed and re-emitted isotropically as a part of the scintillation process. On average scintillation light is delayed with respect to the direct Cherenkov light due to chromatic dispersion and the finite time of the scintillation processes; the early light thus contains directional information.

Using a Geant4~simulation of a spherical detector with radius of 6.5~m and photodetectors with TTS of 0.1~ns, we have shown that for electrons originated at the center a time cut on the early light is effective at isolating the directional light, improving the ratio of Cherenkov light to scintillation light from 0.02 to 0.63. This ratio is degraded by a factor of 2.5 if current photodetectors with $\sim$1~ns resolution were used. This can be improved by factors of 1.6 or 1.4 if more red-sensitive photodetectors or scintillators with narrower emission spectra are used.

Reconstruction algorithms developed for water Cherenkov detectors have been applied to this early light. The algorithms are able to converge on reasonable reconstructed vertices and directions for all simulated energies: 5~MeV, 2.1~MeV, 1.4~MeV. As expected, the reduction in photon statistics with lower energies leads to a broadening of the reconstructed vertex by $\sim$1~cm. Similarly, the mean $\cos(\theta)$ between the reconstructed and initial directions drops from 0.88 to 0.74.
% reconstructed direction drops from 0.88 to 0.74 as the $\cos(\theta)$ distribution broadens.

This technique is promising, and we plan to continue work on the topic. The ability to reconstruct direction in kiloton-scale scintillation detectors would expand capabilities of neutrino experiments, especially those also searching for neutrino-less double-beta decay. More generally, this technique could be applied wherever scintillation-based detectors are used.
